---
title: "Optimising the Cold Water Patch Detection Algorithm"
subtitle: "BSc Project – Progress Report"
author: "Etienne Dürig"
date: "2026-04-01"
format:
  revealjs:
    theme: default
    slide-number: true
---

```{r}
#| include: false

library(terra)
library(sf)
library(dplyr)
library(tidyverse)
library(tictoc)
library(tmap)
library(exactextractr)


step_m            = 500
buffer_px         = 3
delta_C           = 1.0
min_patch_area_m2 = 2
round_to          = 0.1
slab_halfwidth_m  = 60
connect_diagonals = TRUE
union_chunk_size  = 50000L

line_path <- file.path("./data/Centerlines_FINAL/Emme_V01.shp")
ras_path <- file.path("./data/thermal_rasters_FINAL/mean_v01emme.tif")


r <- terra::rast(ras_path)
line <- sf::st_read(line_path, quiet = TRUE) |> sf::st_zm(TRUE, "ZM")

stopifnot(terra::nlyr(r) == 1)
names(r) <- "T"
if (terra::is.lonlat(r)) stop("Raster must be in a metric CRS.")

rres   <- terra::res(r)
cell_m <- mean(rres)

if (!is.na(sf::st_crs(line)) &&
  !identical(terra::crs(r), sf::st_crs(line)$wkt)) {
  line <- sf::st_transform(line, terra::crs(r))
}

g <- sf::st_union(line)

if (inherits(g, "sfc_GEOMETRYCOLLECTION"))
  g <- sf::st_collection_extract(g, "LINESTRING")
if (inherits(g, "sfc_MULTILINESTRING")) {
  g <- sf::st_line_merge(g)
  if (inherits(g, "sfc_GEOMETRYCOLLECTION"))
    g <- sf::st_collection_extract(g, "LINESTRING")
}

rfactor <- if (!is.null(round_to) && round_to > 0) 1 / round_to else NA_real_

# ── 1. BUILD SLABS ────────────────────────────────────────────────────────

L <- as.numeric(sf::st_length(line))
S <- step_m

pos <- sort(unique(c(0, seq(0, 1, by = S / L), 1)))
npt <- length(pos)

if (npt < 2) stop("Too few stations for step = ", S)

from <- pos[1:(npt - 1)]
to   <- pos[2:npt]
k    <- length(from)

g1 <- sf::st_geometry(line)[1]

subs_list <- vector("list", k)
for (j in seq_len(k)) {
  subs_list[[j]] <- lwgeom::st_linesubstring(g1, from[j], to[j])
}

subs <- do.call(c, subs_list)

slabs <- sf::st_buffer(subs,
                       dist        = slab_halfwidth_m,
                       endCapStyle = "FLAT",
                       joinStyle   = "MITRE",
                       mitreLimit  = 2)
slabs_sf <- sf::st_sf(geometry = slabs)

slabs_terra <- terra::vect(slabs_sf)

buffer_m <- cell_m * buffer_px

slabs_buffer <- sf::st_buffer(subs,
                              dist        = buffer_m,
                              endCapStyle = "FLAT",
                              joinStyle   = "MITRE",
                              mitreLimit  = 2)

slaps_buffer_sf <- sf::st_sf(geometry = slabs_buffer) %>% mutate(slap_id = row_number())

slaps_buffer_vect <- slaps_buffer_sf %>%
  mutate(slap_id = row_number()) %>%
  terra::vect()

slabs <- sf::st_buffer(subs,
                       dist        = slab_halfwidth_m,
                       endCapStyle = "FLAT",
                       joinStyle   = "MITRE",
                       mitreLimit  = 2)
slabs_sf <- sf::st_sf(geometry = slabs) %>% mutate(slap_id = row_number())
sf::write_sf(slabs_sf, "slabs_sf.shp")

r_rounded <- terra::rast("r_rounded.tiff")
r <- terra::rast("./data/thermal_rasters_FINAL/mean_v01emme.tif")
slabs_sf_vect <- terra::vect("slabs_sf.shp")
```

## Current CWP algorithm

- Inspected the algorithm
- Speeding Up Algorithm: Profiling into adjusting
- Profiling hard, since `terra::extract` crashed my PC

## Reimplementation along rationale

GDAL and terra feature a set of operations (grammar) for manipulating raster- and vectordata.

::: {style="font-size: 0.65em;"}
| Operation | terra | GDAL |
|---|---|---|
| zonal | `terra::zonal` | `gdal raster zonal` |
| focal | `terra::focal` | `gdal raster neighbours` |
| raster algebra | `A / B * C` | `gdal raster calc` |
| reclassification | `terra::classify` | `gdal raster reclassify` |
| rasterisation | `terra::rasterize` | `gdal vector rasterize` |
| polygonisation | `terra::polygonize` | `gdal raster polygonize` |
:::

## Reimplementation along rationale

Terra is easier to use but GDAL is often considerably faster.

```{r}
#| echo: true

tic("Terra based rasterisation:")
zone_r_big <- terra::rasterize(slabs_sf_vect, r, field = "slap_id")
toc()
```

```{r}
#| echo: true

e <- ext(r)
ext_string <- paste(e$xmin, e$ymin, e$xmax, e$ymax, sep = ",")
resolution <- res(r)
res_string <- paste(resolution, collapse = ",")

tic("GDAL based rasterization:")
system(paste0(
  'gdal vector rasterize -i slabs_sf.shp -o zone_r_big.tiff ',
  '--extent ', ext_string, ' --resolution ', res_string,
  ' --overwrite --ot Int32 -a slap_id --optimization RASTER '
))
toc()
```

## Reimplementation along rationale

- If the algorithm can be reexpressed in terms of typical geodata operations, then it can be expressed in GDAL.
- There might be speedups possible

## Reimplementation along rationale

`exact_extract` is much faster than `terra::extract` for large rasters.

## Speed comparison: Old vs. New code (On HPC)

- New function takes 29 minutes
- Old function around 1h 10 min
- (Better comparison comming up)

## Memory Requirement

- Memory requirements can be a limiting factor too. A node on the HPC features 200 GB RAM and 20 cores
- For 20 Processes: only 10 GB RAM per process available
- Reworked Algorithm can run on 8 GB RAM
- Old Algorithm needs + 10 GB RAM

::: {.callout-note}
Reworked algorithm can be farmed out using the maximal core count per node — resource effective.
:::

## Are the functions equivalent?

- Hard to determine theoretically.
- Trying a view combinations and hoping that they are representative for all combinations
- Two examples with different `buffer width` and `slab length`.
## Parameter Combination 1 – Output comparison

```{r}
#| echo: false
#| message: false

cur_p1 <- sf::st_read(
  "comparison_shapefiles/current_function/param_1/final_polys.shp",
  quiet = TRUE
)
new_p1 <- sf::st_read(
  "comparison_shapefiles/new_function/param_1/final_polys.shp",
  quiet = TRUE
)

tmap::tmap_mode("view")

tmap::tm_shape(cur_p1, name = "Current function") +
  tmap::tm_polygons(
    col          = "red",
    alpha        = 0.4,
    border.col   = "red",
    border.alpha = 0.8
  ) +
tmap::tm_shape(new_p1, name = "New function") +
  tmap::tm_polygons(
    col          = "blue",
    alpha        = 0.4,
    border.col   = "blue",
    border.alpha = 0.8
  )
```

## Parameter Combination 2 – Output comparison

```{r}
#| echo: false
#| message: false

cur_p2 <- sf::st_read(
  "comparison_shapefiles/current_function/param_2/final_polys.shp",
  quiet = TRUE
)
new_p2 <- sf::st_read(
  "comparison_shapefiles/new_function/param_2/final_polys.shp",
  quiet = TRUE
)

tmap::tmap_mode("view")

tmap::tm_shape(cur_p2, name = "Current function") +
  tmap::tm_polygons(
    col          = "red",
    alpha        = 0.4,
    border.col   = "red",
    border.alpha = 0.8
  ) +
tmap::tm_shape(new_p2, name = "New function") +
  tmap::tm_polygons(
    col          = "blue",
    alpha        = 0.4,
    border.col   = "blue",
    border.alpha = 0.8
  )
```
