# Protocoll
```{r}

#install.packages("sensobol")
#install.packages("deSolve")
install.packages("tidyverse")

```


```{r}
library(deSolve)
library(sensobol)
library(tidyverse)

```


```{r}

log_growth <- function(t,N,parms) {
    r <- parms[1]
    K <- parms[2]
    dNdt <- N*r*(1 - (N/K))
    list(n_cells = dNdt)
}


```

```{r}

t <- seq(0, 50, 0.1)
N_start <- 2
K <- 10^4
r <- 0.2

```

```{r}
bacterial_growth <- deSolve::rk(y = N_start,times =t,parms = c(r,K), func = log_growth
)

bacterial_growth %>%
    as_tibble() %>%
    mutate(n_cells = `1`) %>%
    ggplot(aes(x = time, y = n_cells)) + geom_line()

```

### Sensitivity analyis

We might not know K and r with perfect accuracy, but there might be some uncertainty. We can use sensobol to do variance decomposition

```{r}
bacterial_growth <- deSolve::rk(y = N_start,times =t,parms = c(r,K), func = log_growth
)

bacterial_growth %>%
    as_tibble() %>%
    mutate(n_cells = `1`) %>%
    ggplot(aes(x = time, y = n_cells)) + geom_line()





```

```{r}

mat <- sensobol::sobol_matrices(
    matrices = c("A", "B", "AB", "BA"),
    N = 2^11,
    params = c("r", "K"),
    order = "second",
    type = "LHS"
)


nrow(mat)
ncol(mat)
```


```{r}

?rk

mat[, "r"] <- qunif(mat[, "r"],0.17, 0.23)
mat[, "K"] <- qunif(mat[, "K"], 0.5* 10^4, 1.5*10^4)
```

```{r}
mat %>%
    as_tibble() %>%
    ggplot(aes(x = r)) + geom_histogram()


mat %>%
    as_tibble() %>%
    ggplot(aes(x = K)) + geom_histogram()

```


```{r}

N_start <- 2
t <- seq(0, 50, 0.1)

bact_growth_run <- function(r,K) {


    bacterial_growth <- deSolve::rk(y = N_start,times=t, parms = c(r,K), func = log_growth) %>% 
        tail(n = 1) %>% as.data.frame()
return(bacterial_growth)


}



```

```{r}



res <- map2_dfr(mat[,"r"], mat[,"K"], bact_growth_run) %>% as_tibble()



```

```{r}



res  %>% 
    as_tibble() %>%
    mutate(
        cell_count = `1`
    ) %>% select(!`1`)


```

```{r}

sobol_indices(
    matrices = matrices,
    Y = 
)

```

So the computation of the sobol indices is hidden behind this function. And the model has to be passed to the function. So I currently see no real way of applying this function to the cold-water detection algorithm. Because the CWP algorithm is a function with a multidimensional output. -> If I however do not find a way of performing a spatial sensitivity analisis and need to lump the model into a scalar value, then `sensobol`would become an opportunity again. ` sensobol` however is a great tool to create the matrices so that I can then run the model. 


--- 

## Testing multisensi

There is a function called mutlisensi, that claims to implement spatial sensitivity analsis. I looked into this too, but to me it seems like the multi sensi function is more of a wrapper around the generall sensitivity topic + it can not callculate the pixle wise sobol indices in parallel. But it features great tools to boil down the dimensionality of the model output => If the computation gets to expensive, then this could hoste a sollution.


--- 17.03.2026


- Recieved the data. Set up the git repository and deposited all the data
- Then I started analysing the code -> hard to read, so i used claude sonnet 4.6 to get commented out version of the  code 
- Got the high level structure of the code. Identified that this has an inner loop and an outer loop.
- The outerloop runs the cold water detection algorithm over hyperparameters.
- So the whole structure around this is not of interest right now. I need to make this CWP algorithm go fast and not the whole hyperparameter search funcitonality.
- So i tasked claude opus 4.6 to extract the CWP algorithm from the code an write it to a new R script. Also It should set up the loading in of the data.
- Now I will perform timing and generally start dooing optimisation using profiling.


I then hit a brick wall. Bad alloc, some c++ problems under the hood. -> I understand that I have to take a closer look at the individual rasters, their size, do some loading in myself ect. I can not just dive head in. So i inspected all the rasters with gdal.

Terra extract seems to be a very memory hungy step. I think this is because it generateds a dataframe and puts the complete dataframe into memory. And this is just to much memory allocation for my "standart device". On a device with much more memory, this step however might be absolutely no Issue. The question now is, what to do. Should I just go to the HPC imediately ? Because large memory consumption does not realy mean bad performance in a  way. Or should this step be rethought to something that is less memory hungry ? -> 

I think this can actually also bee done with zonal statistics and reclassificaiton ? 


This might be a possibility. This then however might cause issues with having to compute the mean around the buffer. Hmm its a pretty hard problem.

I think the best thing for me to do right now is to get a jupyter notebook, and visually interpret and decompose the code segment by segment. This, because it is actually so complicated, that I am pooking around in a stack of hay on the look for a needle right now.


### March 19 ---

Finished the code that I ran in my jupyter notebook (did that because terra::explode in their function caused bad::alloc) issues for me. My code worked in the jupyter notebook, but now that i wraped everything into a function, i get bad alloc too. Could it be due to memory allocation of R for functions ? I have to look into that right now (I allready used gc() and rm() to help with the garbage collection, but this is of no help appraently). After reseach i found that terraOptions(memfrac) can be used to set the amount of memory that terra is allowed to consume. I increased that to 0.8. Maybe the other function from the Hydro team works to now on my pc. Because this PC has 30 GB of memory, so there is some room and a realistic chance that the computer can handle the memoryspike.

-- The terra memory adjustment has not really worked, it still crashed at the same location. I improved rm() gc() now. If this won't work, then I am going to move to the HPC. Not even this worked, moving to the hpc as I just have to little memory apparently? 


I set up the function on the HPC. Sadly there indeed is a large memory spike over 50 MG. The rest of the function is rather tame at around 10 - 15 GB. I think the spike occurs during the zonal operation of terra. Hoever I have to look deeper into that.


Maybe, since all the data has to be read into memory for the calculation to work ? Don't know. Does not make to much sense to me to be honest since a lot of statistical functions like sum() mean() etc can also be implemented in a "streaming" fassion. -> To do: Try extract_extract. Try to run the other function.

