```{r}

#install.packages("sensobol")
#install.packages("deSolve")
install.packages("tidyverse")

```


```{r}
library(deSolve)
library(sensobol)
library(tidyverse)

```


```{r}

log_growth <- function(t,N,parms) {
    r <- parms[1]
    K <- parms[2]
    dNdt <- N*r*(1 - (N/K))
    list(n_cells = dNdt)
}


```

```{r}

t <- seq(0, 50, 0.1)
N_start <- 2
K <- 10^4
r <- 0.2

```

```{r}
bacterial_growth <- deSolve::rk(y = N_start,times =t,parms = c(r,K), func = log_growth
)

bacterial_growth %>%
    as_tibble() %>%
    mutate(n_cells = `1`) %>%
    ggplot(aes(x = time, y = n_cells)) + geom_line()

```

### Sensitivity analyis

We might not know K and r with perfect accuracy, but there might be some uncertainty. We can use sensobol to do variance decomposition

```{r}
bacterial_growth <- deSolve::rk(y = N_start,times =t,parms = c(r,K), func = log_growth
)

bacterial_growth %>%
    as_tibble() %>%
    mutate(n_cells = `1`) %>%
    ggplot(aes(x = time, y = n_cells)) + geom_line()





```

```{r}

mat <- sensobol::sobol_matrices(
    matrices = c("A", "B", "AB", "BA"),
    N = 2^11,
    params = c("r", "K"),
    order = "second",
    type = "LHS"
)


nrow(mat)
ncol(mat)
```


```{r}

?rk

mat[, "r"] <- qunif(mat[, "r"],0.17, 0.23)
mat[, "K"] <- qunif(mat[, "K"], 0.5* 10^4, 1.5*10^4)
```

```{r}
mat %>%
    as_tibble() %>%
    ggplot(aes(x = r)) + geom_histogram()


mat %>%
    as_tibble() %>%
    ggplot(aes(x = K)) + geom_histogram()

```


```{r}

N_start <- 2
t <- seq(0, 50, 0.1)

bact_growth_run <- function(r,K) {


    bacterial_growth <- deSolve::rk(y = N_start,times=t, parms = c(r,K), func = log_growth) %>% 
        tail(n = 1) %>% as.data.frame()
return(bacterial_growth)


}



```

```{r}



res <- map2_dfr(mat[,"r"], mat[,"K"], bact_growth_run) %>% as_tibble()



```

```{r}



res  %>% 
    as_tibble() %>%
    mutate(
        cell_count = `1`
    ) %>% select(!`1`)


```

```{r}

sobol_indices(
    matrices = matrices,
    Y = 
)

```

So the computation of the sobol indices is hidden behind this function. And the model has to be passed to the function. So I currently see no real way of applying this function to the cold-water detection algorithm. Because the CWP algorithm is a function with a multidimensional output. -> If I however do not find a way of performing a spatial sensitivity analisis and need to lump the model into a scalar value, then `sensobol`would become an opportunity again. ` sensobol` however is a great tool to create the matrices so that I can then run the model. 


--- 

## Testing multisensi

There is a function called mutlisensi, that claims to implement spatial sensitivity analsis. I looked into this too, but to me it seems like the multi sensi function is more of a wrapper around the generall sensitivity topic + it can not callculate the pixle wise sobol indices in parallel. But it features great tools to boil down the dimensionality of the model output => If the computation gets to expensive, then this could hoste a sollution.

